% =====================================================================
%  Chapter 1: Probability, Statistics, and the Language of Uncertainty
%  Source: Review.tex (Overleaf) + P1 transcription
% =====================================================================

\chapter{Probability, Statistics, and the Language of Uncertainty}
\label{chap_Review}

\section{A Paradigm Shift: From Classical to Frequentist to Bayesian}
\label{sec_paradigm}

``Probability approach in econometrics'' by Trygve \citet{Haavelmo1944} stands
at the beginning of a paradigm shift in statistical thinking; the
\emph{classical} interpretation of statistics is giving way to the
\emph{frequentist} interpretation. Under the classical interpretation,
probabilities are linked to the principle of indifference and statistical
events are governed by natural symmetry (e.g.\ coin tosses, die rolls).

Under the frequentist interpretation, probabilities of events are linked to
the frequency with which a given event is observed:
\begin{equation}
  P(x) = \lim_{n_t \to \infty} \frac{n_x}{n_t},
\end{equation}
where $n_x$ is the number of times event $x$ is observed and $n_t$ is the
total number of trials.

This probabilistic approach to statistics, which encompasses Bayesian
statistics\index{Bayesian statistics}, represents the world in terms of
``degrees of belief'' (Bayesian priors). More formally, \textbf{Bayes'
Theorem}\index{Bayes' Theorem} is expressed as the conditional probability of
event $A$ given $B$:

\begin{tcolorbox}[keyresult]
\begin{equation}
  P(A \mid B) = \frac{P(B \mid A)\,P(A)}{P(B)}
\end{equation}
\end{tcolorbox}


\section{Random Variables: Types and Typology}
\label{sec_RVtypes}

A random variable, usually written as $X$, is a variable whose possible values
are numerical outcomes of a random phenomenon. A random variable $X$ is a
(real) number with the following characteristics:

\begin{enumerate}[label=\alph*)]
  \item The value of $X$ is unknown.
  \item The possible values $X$ can take are known; the set of such values is
    called the \emph{support}, denoted $\boldsymbol{\mathcal{S}}_X$.
  \item If $\boldsymbol{\mathcal{A}} \subseteq \mathbb{R}$, we know the
    probability that $X$ belongs to $\boldsymbol{\mathcal{A}}$, denoted
    $P(X \in \boldsymbol{\mathcal{A}})$.
  \item If the value $x$ of $X$ becomes known such that $x=X$, then $x$ is
    called a \emph{realisation} of $X$.
\end{enumerate}

Random variables fall into two categories:

\begin{itemize}
  \item \textbf{Discrete random variables}\index{random variable!discrete}:
    Take only a countable number of distinct values (0, 1, 2, \ldots).
    Usually (but not necessarily) counts.
  \item \textbf{Continuous random variables}\index{random variable!continuous}:
    Take an infinite number of possible values. Usually measurements (height,
    weight, time).
\end{itemize}

We can also distinguish \emph{qualitative} (categorical) from
\emph{quantitative} variables:

\begin{itemize}
  \item \textbf{Nominal variables}\index{random variable!nominal}: Mutually
    exclusive, unordered categories (e.g.\ suburb type).
  \item \textbf{Ordinal variables}\index{random variable!ordinal}: Ordered
    categories where the differences between values are not meaningful (e.g.\
    quality-of-life rankings on a 1--10 scale).
  \item \textbf{Interval variables}\index{random variable!interval}:
    Differences are meaningful but there is no true zero (e.g.\ temperature
    in Celsius).
  \item \textbf{Ratio variables}\index{random variable!ratio}: A true zero
    exists and ratios are meaningful (e.g.\ height, weight, income).
\end{itemize}

\begin{table}[htp]
\centering
\caption{Basic statistical operations by variable type}
\label{tbl_Vars}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}lcccc@{}}
\toprule
& \multicolumn{4}{c}{\emph{Variable type}} \\
\cmidrule{2-5}
\emph{Operation} & Nominal & Ordinal & Interval & Ratio \\
\midrule
Frequency distribution    & Yes & Yes & Yes & Yes \\
Median and percentiles    & No  & Yes & Yes & Yes \\
Addition / subtraction    & No  & No  & Yes & Yes \\
Mean, s.d., s.e.          & No  & No  & Yes & Yes \\
Ratio / coefficient of variation & No & No & No & Yes \\
\bottomrule
\end{tabular}
\renewcommand{\arraystretch}{1}
\end{table}


\section{Expectation and Variance}
\label{sec_Expectation}

\subsection{Expectation of a Discrete Random Variable}

\begin{equation}
  \E(X) = x_1 p_1 + x_2 p_2 + \cdots + x_n p_n = \sum_{i=1}^{n} x_i p_i
\end{equation}

In a finite sample, $\E(X)$ denotes the sample mean, and the sum of all
probabilities must equal one: $\sum_i p_i = 1$.

\paragraph{Example: The St.\ Petersburg Paradox}\index{St. Petersburg paradox}

Consider a gamble where an unbiased coin is tossed until tails ($T$) appears.
If $T$ shows up on the $n$-th toss, the payoff is $\$2^{n-1}$. How much would
you be willing to wager to participate?

\begin{align*}
  \E(X) &= \$2^0 \cdot \tfrac{1}{2} + \$2^1 \cdot \tfrac{1}{4}
           + \$2^2 \cdot \tfrac{1}{8} + \cdots \\
        &= \$1\cdot\tfrac{1}{2} + \$2\cdot\tfrac{1}{4}
           + \$4\cdot\tfrac{1}{8} + \cdots
         = \sum_{i=1}^{\infty} \$0.50 = \$\infty
\end{align*}

This is the St.~Petersburg Paradox, famously described by Daniel
Bernoulli\index{Bernoulli, Daniel} (1700--1782) to illustrate the failure of
the expected value principle in decision theory under uncertainty. The expected
payout is infinite, yet willingness to pay (WTP)\index{willingness to pay}
must be much less. The classical resolution introduces an explicit utility
function and the presumption of diminishing marginal utility, implying
$\E[U] < \E[X]$.

\paragraph{Example: Expected value of a pair of dice}

\begin{equation*}
  \E(X) = \tfrac{1}{6}\cdot1 + \tfrac{1}{6}\cdot2 + \cdots
          + \tfrac{1}{6}\cdot6 = 3.5
\end{equation*}

For a pair of dice, $\mu_X = 7$. Bivariate probabilities are shown in
Table~\ref{tbl_Dice}.

\begin{table}[htp]
\centering
\caption{Bivariate probabilities for a pair of dice}
\label{tbl_Dice}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}lll|lll@{}}
\toprule
$X$ & $p$ & $Xp$ & $X$ & $p$ & $Xp$ \\
\midrule
$x_1$ & $p_1$ & $x_1p_1$ & 2  & 1/36 & 2/36  \\
$x_2$ & $p_2$ & $x_2p_2$ & 3  & 2/36 & 6/36  \\
      &       &           & 4  & 3/36 & 12/36 \\
      &       &           & 5  & 4/36 & 20/36 \\
      &       &           & 6  & 5/36 & 30/36 \\
      &       &           & 7  & 6/36 & 42/36 \\
      &       &           & 8  & 5/36 & 40/36 \\
      &       &           & 9  & 4/36 & 36/36 \\
      &       &           & 10 & 3/36 & 30/36 \\
      &       &           & 11 & 2/36 & 22/36 \\
$x_n$ & $p_n$ & $x_np_n$ & 12 & 1/36 & 6/36  \\
\midrule
\multicolumn{3}{l}{$\E(X)=\sum x_i p_i$} &&
\multicolumn{2}{r}{252/36 = 7} \\
\bottomrule
\end{tabular}
\renewcommand{\arraystretch}{1}
\end{table}

\subsection{Expectation Rules}

\begin{enumerate}
  \item $\E[X + Y + Z] = \E[X] + \E[Y] + \E[Z]$
  \item $\E[bX] = b\E[X]$
  \item $\E[b] = b$
  \item $\E[a + bX] = a + b\E[X]$
\end{enumerate}

\paragraph{Example} Let $V \sim \mathcal{N}(0,1)$ and $Z \equiv \mu + \sigma V$.
Then $\E[Z] = \mu$ and $\Var[Z] = \sigma^2$, since $\E[V]=0$ and
$\E[V^2]=1$.

\subsection{Variance}

\begin{equation}
  \Var(X) = \sigma^2_X = \E\bigl[(X-\mu_X)^2\bigr]
  = \sum_{i=1}^{n}(x_i - \mu_X)^2 p_i
\end{equation}

Alternative expression (computing formula):
\begin{align}
  \Var(X) &= \E\bigl[X^2 - 2\mu_X X + \mu_X^2\bigr] \nonumber \\
          &= \E[X^2] - 2\mu_X \E[X] + \mu_X^2 \nonumber \\
          &= \E[X^2] - \mu_X^2
\end{align}

\subsection{Fixed and Random Decomposition of Variables}

Let the random variable $X$ be decomposed into a constant mean $\mu_X$ and a
stochastic disturbance term $u$:
\begin{equation}
  X = \mu_X + u, \qquad \E[u] = \E[X - \mu_X] = 0
\end{equation}
Since all variation in $X$ is due to $u$, the variance of $X$ equals the
variance of the disturbance:
\begin{equation}
  \sigma^2_X = \E[(X-\mu_X)^2] = \E[u^2] = \sigma^2_u
\end{equation}
This decomposition into a \emph{systematic} (fixed) component and a
\emph{stochastic} (random) component is the conceptual foundation of every
regression model in this book.


\section{Continuous Random Variables}
\label{sec_ContRV}

All random variables have a \textbf{probability density function}
(PDF)\index{density function!probability} and a \textbf{cumulative density
function} (CDF)\index{density function!cumulative}:
\begin{equation}
  \Pr[A \leq x \leq B] = \int_A^B f(x)\,dx, \qquad
  F(x) = \int_{-\infty}^{x} f(t)\,dt = F(B) - F(A)
\end{equation}

The expectation and variance of a continuous random variable:
\begin{equation}
  \E[X] = \int_{-\infty}^{\infty} x\,f(x)\,dx = \mu_x,
  \qquad
  \Var(X) = \int_{-\infty}^{\infty}(x - \mu_x)^2 f(x)\,dx = \sigma_x^2
\end{equation}

\begin{figure}[htp]
\centering
\begin{tikzpicture}[>=Stealth, scale=1.1]
  % Left panel: PDF
  \begin{scope}[xshift=0cm]
    \draw[->] (-2.5,0) -- (2.7,0) node[right] {$x$};
    \draw[->] (0,-0.1) -- (0,1.6) node[above] {$f(x)$};
    \draw[very thick, domain=-2.4:2.4, samples=80]
      plot (\x, {1.3*exp(-0.8*\x*\x)});
    % Shaded region between -1.5 and -0.5
    \fill[gray!30, domain=-1.5:-0.5, samples=40]
      (-1.5,0) -- plot (\x, {1.3*exp(-0.8*\x*\x)}) -- (-0.5,0) -- cycle;
    \draw[dashed] (-1.5,0) node[below]{\small$A$}
                         -- (-1.5, {1.3*exp(-0.8*1.5*1.5)});
    \draw[dashed] (-0.5,0) node[below]{\small$B$}
                         -- (-0.5, {1.3*exp(-0.8*0.5*0.5)});
    \node[below=18pt] at (0,0) {PDF:\; $f(x) \equiv F'(x)$};
  \end{scope}
  % Right panel: CDF
  \begin{scope}[xshift=6cm]
    \draw[->] (-2.5,0) -- (2.7,0) node[right] {$x$};
    \draw[->] (0,-0.1) -- (0,1.6) node[above] {$F(x)$};
    \draw[very thick, domain=-2.4:2.4, samples=80]
      plot (\x, {1.3/(1+exp(-2.2*\x))});
    \draw[dashed] (-1.5,0) node[below]{\small$A$}
                         -- (-1.5, {1.3/(1+exp(2.2*1.5))});
    \draw[dashed] (-0.5,0) node[below]{\small$B$}
                         -- (-0.5, {1.3/(1+exp(2.2*0.5))});
    \node[below=18pt] at (0,0) {CDF:\; $F(x) = \int_{-\infty}^x f(t)\,dt$};
  \end{scope}
\end{tikzpicture}
\caption{Probability density function (PDF) and cumulative density function
(CDF) for a continuous random variable. The shaded area in the PDF equals
$\Pr[A \leq x \leq B]$.}
\label{fig_PDFCDF}
\end{figure}


\section{Moments: Skewness and Kurtosis}
\label{sec_Moments}

Higher-order moments describe the shape of a distribution beyond its mean and
variance.

\noindent\textbf{Skewness} ($3$rd central moment / $\sigma^3$):
\begin{equation}
  \gamma_1 = \frac{\E[(X-\mu)^3]}{\sigma^3}
  \qquad
  \begin{cases} > 0 & \text{right-skewed (positive tail)} \\
                = 0 & \text{symmetric} \\
                < 0 & \text{left-skewed (negative tail)}
  \end{cases}
\end{equation}

\noindent\textbf{Kurtosis} ($4$th central moment / $\sigma^4$):
\begin{equation}
  \gamma_2 = \frac{\E[(X-\mu)^4]}{\sigma^4}
  \qquad
  \begin{cases} > 3 & \text{leptokurtic (fat tails)} \\
                = 3 & \text{mesokurtic (normal)} \\
                < 3 & \text{platykurtic (thin tails)}
  \end{cases}
\end{equation}


\section{Covariance, Correlation, and Joint Distributions}
\label{sec_CovCorr}

\begin{equation}
  \Cov(X,Y) = \sigma_{XY} = \E[(X-\mu_X)(Y-\mu_Y)]
             = \E[XY] - \mu_X\mu_Y
\end{equation}

\begin{equation}
  \rho_{XY} = \frac{\Cov(X,Y)}{\sigma_X \sigma_Y} \;\in\; [-1,+1]
\end{equation}

\noindent\textbf{Covariance and correlation rules:}
\begin{enumerate}[label=(\roman*)]
  \item $\Cov(X,X) = \Var(X)$
  \item $\Cov(aX,bY) = ab\,\Cov(X,Y)$
  \item $\Var(X+Y) = \Var(X) + \Var(Y) + 2\Cov(X,Y)$
  \item If $X \ind Y$: $\Cov(X,Y) = 0$ and $\Var(X+Y) = \Var(X)+\Var(Y)$
\end{enumerate}

\noindent\textbf{Sample estimators:}
\begin{equation}
  \hat{\sigma}_{XY} = \frac{1}{n-1}\sum_{i=1}^{n}(x_i-\bar{x})(y_i-\bar{y}),
  \qquad
  \hat{\rho}_{XY} = \frac{\hat{\sigma}_{XY}}{\hat{\sigma}_X\hat{\sigma}_Y}
\end{equation}


\section{Estimator Properties}
\label{sec_EstimatorProps}

Let $\hat{\theta}$ be an estimator of population parameter $\theta$.

\subsection{Unbiasedness}

$\hat{\theta}$ is \textbf{unbiased} if $\E[\hat{\theta}] = \theta$.
The \textbf{bias} is $\text{Bias}(\hat{\theta}) = \E[\hat{\theta}] - \theta$.

\subsection{Efficiency}

Among all unbiased estimators, $\hat{\theta}$ is \textbf{efficient} (best) if
it has the smallest variance. The Cramér-Rao lower bound defines the minimum
variance achievable by an unbiased estimator.

\begin{figure}[htp]
\centering
\begin{tikzpicture}[>=Stealth, scale=1.0]
  % Efficient estimator (narrow)
  \draw[very thick, domain=-2:2, samples=60]
    plot (\x, {1.8*exp(-3*\x*\x)}) node[right]{\small Efficient $\hat\theta_1$};
  % Inefficient estimator (wide)
  \draw[thick, dashed, domain=-3.5:3.5, samples=80]
    plot (\x, {0.9*exp(-0.6*\x*\x)}) node[right]{\small Inefficient $\hat\theta_2$};
  % True value
  \draw[gray, thin] (0,-0.1) -- (0,1.95);
  \node[below] at (0,-0.1) {$\theta$};
  \draw[->] (-3.8,0) -- (4.2,0) node[right] {$\hat\theta$};
  \draw[->] (0,-0.15) -- (0,2.2);
\end{tikzpicture}
\caption{Efficiency: $\hat\theta_1$ is more efficient than $\hat\theta_2$
because it has smaller variance around the true parameter value $\theta$.}
\label{fig_Efficiency}
\end{figure}

\subsection{Mean Squared Error}

\begin{tcolorbox}[keyresult]
\begin{equation}
  \MSE(\hat{\theta}) = \E[(\hat{\theta} - \theta)^2]
  = \Var(\hat{\theta}) + \bigl[\text{Bias}(\hat{\theta})\bigr]^2
\end{equation}
\end{tcolorbox}

This decomposition shows the fundamental bias--variance trade-off: a slightly
biased estimator may have lower MSE than an unbiased one if its variance is
sufficiently smaller.

\subsection{Consistency}

$\hat{\theta}_n$ is \textbf{consistent} if $\hat{\theta}_n
\xrightarrow{p} \theta$ as $n \to \infty$:
\begin{equation}
  \lim_{n \to \infty} P(|\hat{\theta}_n - \theta| > \varepsilon) = 0
  \quad \forall\,\varepsilon > 0
\end{equation}

\begin{figure}[htp]
\centering
\begin{tikzpicture}[>=Stealth, scale=0.95]
  \def\tp{0}  % true parameter
  \foreach \s/\lbl in {1.2/{$n=10$}, 0.7/{$n=50$}, 0.3/{$n=200$}}{
    \draw[domain={\tp-3.5*\s}:{\tp+3.5*\s}, samples=60, thick]
      plot (\x, {1/(\s*sqrt(2*3.14159))*exp(-0.5*((\x-\tp)/\s)^2)});
  }
  \draw[->] (-4.2,0) -- (4.2,0) node[right]{$\hat\theta$};
  \draw[->] (0,-0.1) -- (0,1.5);
  \draw[gray, thin] (\tp,-0.05) -- (\tp,1.45);
  \node[below] at (\tp,-0.05) {$\theta$};
  \node[right] at (1.5,1.3) {\small $n=200$};
  \node[right] at (1.8,0.55) {\small $n=50$};
  \node[right] at (2.5,0.28) {\small $n=10$};
\end{tikzpicture}
\caption{Consistency: as $n$ increases, the sampling distribution of
$\hat\theta$ shrinks around the true value $\theta$.}
\label{fig_Consistency}
\end{figure}


\section{Statistical Inference}
\label{sec_Inference}

\subsection{The Hypothesis Testing Framework}

\begin{enumerate}
  \item Formulate $H_0$ (null) and $H_1$ (alternative)
  \item Choose significance level $\alpha$ (typically 0.01, 0.05, 0.10)
  \item Compute test statistic and $p$-value
  \item Reject $H_0$ if $p < \alpha$
\end{enumerate}

\begin{table}[htp]
\centering
\caption{Decision outcomes in hypothesis testing}
\label{tbl_TypeErrors}
\renewcommand{\arraystretch}{1.3}
\begin{tabular}{@{}lcc@{}}
\toprule
& $H_0$ True & $H_0$ False \\
\midrule
Fail to reject $H_0$ & Correct ($1-\alpha$) & Type II error ($\beta$) \\
Reject $H_0$         & Type I error ($\alpha$) & Correct ($1-\beta$) \\
\bottomrule
\end{tabular}
\renewcommand{\arraystretch}{1}
\end{table}

\subsection{The $z$-Statistic and $p$-Values}

For a test of $H_0: \mu = \mu_0$:
\begin{equation}
  z = \frac{\bar{x} - \mu_0}{\sigma/\sqrt{n}}
  \sim \mathcal{N}(0,1) \quad \text{under } H_0
\end{equation}

\begin{tcolorbox}[keyresult]
Two-sided $p$-value:\quad $p(\hat{z}) = 2\bigl(1 - \Phi(|\hat{z}|)\bigr)$
\end{tcolorbox}

\subsection{Statistical Power}

The \textbf{power} of a test is $1 - \beta$, the probability of correctly
rejecting a false null hypothesis.

\begin{figure}[htp]
\centering
\begin{tikzpicture}[>=Stealth, scale=1.0]
  % H0 distribution (centred at 0)
  \draw[very thick, domain=-3:3, samples=80]
    plot (\x, {1.2*exp(-0.8*\x*\x)});
  \node at (-1.8, 0.9) {$H_0$};
  % H1 distribution (centred at 2)
  \draw[very thick, dashed, domain=-1:5, samples=80]
    plot (\x, {1.2*exp(-0.8*(\x-2)*(\x-2))});
  \node at (3.8, 0.9) {$H_1$};
  % Critical value
  \draw[gray, thin] (1.4,-0.05) -- (1.4,1.3);
  \node[below] at (1.4,-0.05) {\small $z_\alpha$};
  % Type II error region (shaded)
  \fill[gray!30, domain=-1:1.4, samples=40]
    (-1,0) -- plot (\x, {1.2*exp(-0.8*(\x-2)*(\x-2))}) -- (1.4,0) -- cycle;
  \node at (0.2, 0.18) {\small $\beta$};
  % Rejection region label
  \node at (2.5, 0.2) {\small Power $= 1-\beta$};
  \draw[->] (-3.5,0) -- (5.5,0) node[right]{$z$};
  \draw[->] (0,-0.1) -- (0,1.5);
\end{tikzpicture}
\caption{Statistical power. The shaded region represents the Type II error
probability $\beta$; power is $1-\beta$.}
\label{fig_Power}
\end{figure}

\subsection{Confidence Intervals}

A $(1-\alpha)\times100\%$ confidence interval for $\mu$:
\begin{equation}
  \bar{x} \pm z_{\alpha/2}\,\frac{\sigma}{\sqrt{n}}
\end{equation}

With unknown $\sigma$, replace $z_{\alpha/2}$ with $t_{\alpha/2, n-1}$.
The \textbf{$t$-distribution}\index{t-distribution} with $\nu$ degrees of
freedom has heavier tails than the standard normal and converges to
$\mathcal{N}(0,1)$ as $\nu \to \infty$.

\subsection{Probability Limits and Asymptotic Theory}

The \textbf{probability limit}\index{probability limit} ($\plim$) satisfies
the following rules:

\begin{enumerate}[label=(\roman*)]
  \item $\plim(X + Y) = \plim(X) + \plim(Y)$
  \item $\plim(XY) = \plim(X)\cdot\plim(Y)$
  \item $\plim(X/Y) = \plim(X)/\plim(Y)$ \;(if $\plim(Y) \neq 0$)
  \item $\plim[g(X)] = g(\plim X)$ \;(continuous $g$; Slutsky's theorem)
  \item $\plim(\bar{X}) = \mu$ \;(law of large numbers)
  \item $\plim(\hat{\sigma}^2) = \sigma^2$ \;(consistency of sample variance)
\end{enumerate}


\section{The Central Limit Theorem}
\label{sec_CLT}

\subsection{Lindberg-Lévy CLT}

For i.i.d.\ random variables $x_t$ with mean $\mu$ and variance $\sigma^2$,
the \textbf{Central Limit Theorem}\index{central limit theorem} (CLT) states
that:

\begin{tcolorbox}[keyresult]
\begin{equation}
  z_t \equiv \frac{1}{\sqrt{n}}\sum_{t=1}^{n}\frac{x_t - \mu}{\sigma}
  \xrightarrow{d} \mathcal{N}(0,1) \quad \text{as } n \to \infty
  \label{eqn_CLT}
\end{equation}
\end{tcolorbox}

\subsection{CLT in Practice: How Large Does $n$ Need to Be?}

The CLT becomes effective at surprisingly small $n$ for many distributions:

\begin{itemize}
  \item \textbf{Binomial:} Approximation improves dramatically at $n=50$.
  \item \textbf{Exponential:} CLT takes effect visibly at $n=50$.
  \item \textbf{$t$-distribution:} Even at $n=10$, centered random variables
    approximate $\mathcal{N}(0,1)$ well (the two distributions are similar
    except for heavier tails in the $t$).
  \item \textbf{Gamma/Beta:} Reasonable approximation at smallest $n$,
    reliable at $n=50$.
  \item \textbf{Log-normal:} CLT kicks in only at $n > 100$ due to high skew.
\end{itemize}

\noindent\textbf{Variance and sample size as substitutes.} The $\chi^2(m)$
distribution approaches $\mathcal{N}(m, 2m)$ as $m$ becomes large. A higher
variance implies a larger denominator in the centering process, which in turn
requires \emph{fewer} observations for the distribution to be approximately
standard normal. Variance and sample size are substitutes in the centering
process.

\renewcommand\bibname{Further reading}
\begin{subappendices}
\bibliographystyle{econometrica}
\bibliography{Literature/OverLord}
\IfFileExists{Literature/Review.bbl}{\input{Literature/Review.bbl}}{\typeout{Need to run bibtex on Review}}
\end{subappendices}
